Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-24 manuscript/manuscript.tex
update ms. (add refs and format.args)

\abstract{%
Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition. However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration. The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data. Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states. One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches. To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis. We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferntial statistics via Bayesian and frequentist multi-level regression modelling. We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences. We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work. Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
}

@@

From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
Indeed, hundreds of years ago, analytical methods were developed to predict time to death (REFs).
The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
Here, in the paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.

@@

In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (REFs) and tutorials (Allison, 2010) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.

@@

Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A). The discrete-time hazard function gives you for each bin the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins. This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins.

{[}{[}this paragraph is not so clear yet\ldots{} Perhaps add error bars to the functions in Figure 2?{]}{]}
The survivor function can help to qualify or provide context to the interpretation of the hazard function. For example, it can give a sense of how many trials contribute to each part of the hazard distribution. If a participant completes 100 trials in an experiment, and the survivor function reaches a probability of 0.03 at the end of timebin (400,500{]}, then only 3\% of trials remain beyond this point, which in this case would amount to 3 trials. Therefore, the error bars in later parts of the hazard function would be wider and less precise compared to earlier parts.

@@

In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer. For example, a task may involve choosing between two response options with only one of them being correct. For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis. Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977). We refer to this extended analysis for choice RT data as EHA/SAT.

@@

The next step is to set up the life table -- custom function setup\_lt() --, calculate the conditional accuracies -- custom function calc\_ca() --, and then plot the descriptive statistics using our custom function plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:

@@

Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009), we relax all three assumptions in model 4. We thus add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.

@@

\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:int-table}Point and 95\% highest posterior density interval estimates, and hazard ratios.}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:int-table} continued}}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endhead
240.00 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82\\
280.00 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03\\
320.00 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45\\
360.00 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91\\
400.00 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.71\\
440.00 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.64\\
480.00 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62\\
520.00 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.62\\
560.00 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.60\\
600.00 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.52\\
240.00 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63\\
280.00 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.13\\
320.00 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90\\
360.00 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99\\
400.00 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68\\
440.00 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58\\
480.00 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58\\
520.00 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.64\\
560.00 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71\\
600.00 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.78\\
240.00 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02\\
280.00 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.66\\
320.00 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.52\\
360.00 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68\\
400.00 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.40\\
440.00 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28\\
480.00 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23\\
520.00 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20\\
560.00 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18\\
600.00 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15\\
240.00 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10\\
280.00 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72\\
320.00 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58\\
360.00 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.34\\
400.00 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.30\\
440.00 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.34\\
480.00 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44\\
520.00 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.58\\
560.00 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65\\
600.00 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.56\\
240.00 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68\\
280.00 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88\\
320.00 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78\\
360.00 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.40\\
400.00 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.33\\
440.00 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.39\\
480.00 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.56\\
520.00 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87\\
560.00 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27\\
600.00 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52\\
240.00 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07\\
280.00 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.94\\
320.00 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.72\\
360.00 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.35\\
400.00 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28\\
440.00 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.35\\
480.00 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55\\
520.00 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00\\
560.00 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.79\\
600.00 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

@@

In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}. Due to practice, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).

@@

240.00 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24,618,458.61\\
280.00 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4,846.09\\
320.00 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.79\\
360.00 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.89\\
400.00 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.19\\
440.00 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.31\\
480.00 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.41\\
520.00 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.40\\
560.00 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.12\\
600.00 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.00\\
240.00 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1,899,836.02\\
280.00 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1,961.83\\
320.00 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.87\\
360.00 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.57\\
400.00 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.14\\
440.00 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.06\\
480.00 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.10\\
520.00 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.00\\
560.00 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.00\\
600.00 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.00\\
240.00 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10,001,085.39\\
280.00 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5,124.44\\
320.00 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.12\\
360.00 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.82\\
400.00 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.58\\
440.00 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.82\\
480.00 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.08\\
520.00 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.00\\
560.00 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.00\\
600.00 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.00\\
240.00 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.00\\
280.00 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.00\\
320.00 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.02\\
360.00 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.38\\
400.00 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.59\\
440.00 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.59\\
480.00 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.30\\
520.00 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.04\\
560.00 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3,282,435.93\\
600.00 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46,319,712,352,328.76\\
240.00 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.00\\
280.00 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.00\\
320.00 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.04\\
360.00 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.59\\
400.00 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.10\\
440.00 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.68\\
480.00 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.20\\
520.00 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3,307.44\\
560.00 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164,758,701.84\\
600.00 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10,165,856,639,901,592.00\\
240.00 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.00\\
280.00 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.00\\
320.00 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.06\\
360.00 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.87\\
400.00 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.57\\
440.00 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.21\\
480.00 & i1500 & 5.52 & -13.87 & 26.57 & 0.95 & 249.51\\

>>>>

\abstract{%
Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition. However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration. The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data. Moreover, well-established analytical approaches, such as event history analysis, are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the timeline of psychological states. One barrier to wider use of event history analysis, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches. To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time event history analysis. We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multi-level regression modelling. We touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences. We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future durections of this work. Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are avilable as .html files to widen access beyond R-users.
}

@@

From a historical perspective, it is worth noting that the development of analytical tools that can estimate or predict whether and when events will occur is not a new endeavour.
Indeed, hundreds of years ago, analytical methods were developed to predict time to death (William Matthew Makeham, 1860).
The same logic has been applied to psychological time-to-event data, as previously demonstrated (Panis et al., 2020).
Here, in the paper, we hope to show the value of event history analysis for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience, as well as provide practical tutorials that provide step-by-step code and instructions in the hope that we can enable others to use event history analysis in a more routine, efficient and effective manner.

@@

In summary, even though event history analysis is a widely used statistical tool and there already exist many excellent reviews (e.g., Blossfeld \& Rohwer, 2002; Box-Steffensmeier, 2004; Hosmer, Lemeshow, \& May, 2011; Teachman, 1983) and tutorials (e.g., Allison, 2010; Landes, Engelhardt, \& Pelletier, 2020) on its general use-cases, we are not aware of any tutorials that are aimed at psychological time-to-event data, and which provide worked examples of the key data processing and multi-level regression modelling steps.
Thereofore, our ultimate goal is twofold: first, we want to convince readers of the many benefits of using event history analysis when dealing with time-to-event data with a focus on psychological time-to-event data, and second, we want to provide a set of practical tutorials, which provide step-by-step instructions on how you actually perform a discrete-time event history analysis on time-to-event data such as RT data, as well as a complementary discrete-time SAT analysis on timed accuracy data.

@@

Discrete-time EHA focuses on the discrete-time hazard function of event occurrence and the discrete-time survivor function (Figure 2). The equations that define both of these functions are reported in the supplementary material (equations 1 and 2 in part A). The discrete-time hazard function gives you for each bin the probability that the event occurs (sometime) in bin t, given that the event does not occur in previous bins. In other words, it reflects the instantaneous likelihood that the event occurs in the current bin, given that it has not yet occurred in the past, i.e., in one of the prior bins. This conditionality in the definition of hazard is what makes the hazard function so diagnostic for studying event occurrence, as an event can physically not occur when it has already occurred before. In contrast, the discrete-time survivor function cumulates the bin-by-bin risks of event \emph{non}occurrence to obtain the probability that the event occurs after bin t. In other words, the survivor function reflects the likelihood that the event occurs in the future, i.e., in one of the subsequent timebins. As explained in part A of the supplementary material, the survivor function provides a context for interpreting the hazard function.

@@

In the context of experimental psychology, it is common for participants to be presented with either a 1-button detection task or a 2-button discrimination task, i.e., a task that has a right and a wrong answer. For example, a task may involve choosing between two response options with only one of them being correct. For such two-choice RT data, the discrete-time EHA can be extended with a discrete-time SAT analysis. Specifically, the hazard function of event occurrence can be extended with the discrete-time conditional accuracy function (see equation 5 in part A of the supplementary material), which gives you the probability that a response is correct given that it is emitted in time bin t (Allison, 2010; Kantowitz \& Pachella, 2021; Wickelgren, 1977). We refer to this extended analysis for choice RT data as EHA/SAT. As explained in part A of the supplementary material, the probability mass function provides a context for interpreting the conditional accuracy function.

@@

The next step is to set up the life table using our custom function setup\_lt(), calculate the conditional accuracies using our custom function calc\_ca(), add the ca(t) estimates to the life table using our custom function join\_lt\_ca(), and then plot the descriptive statistics using our custom function plot\_eha(). When creating the plots, some warning messages will likely be generated, like these:

@@

Based on previous work (Panis, 2020; Panis, Moran, et al., 2020; Panis \& Schmidt, 2022; Panis et al., 2017; Panis \& Wagemans, 2009) we expect nonlinear effects of trial number. We thus relax all three assumptions in model 4. We add a squared term for the continuous predictor centered trial number -- I(trial\_c\^{}2) -- and include interaction terms.

@@

\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:int-table}Point and 95\% highest posterior density interval estimates, and hazard ratios.}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:int-table} continued}}\\
\toprule
bin\_endpoint & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{mean} & \multicolumn{1}{c}{.lower} & \multicolumn{1}{c}{.upper} & \multicolumn{1}{c}{.width} & \multicolumn{1}{c}{hazard ratio}\\
\midrule
\endhead
240 & c500 & 2.18 & 0.33 & 4.05 & 0.95 & 8.82080\\
280 & c500 & 1.11 & -0.02 & 2.11 & 0.95 & 3.03199\\
320 & c500 & 0.37 & -0.34 & 1.04 & 0.95 & 1.45383\\
360 & c500 & -0.09 & -0.70 & 0.48 & 0.95 & 0.91397\\
400 & c500 & -0.35 & -1.02 & 0.34 & 0.95 & 0.70802\\
440 & c500 & -0.45 & -1.50 & 0.56 & 0.95 & 0.63522\\
480 & c500 & -0.48 & -2.32 & 1.27 & 0.95 & 0.62035\\
520 & c500 & -0.48 & -3.57 & 2.52 & 0.95 & 0.61980\\
560 & c500 & -0.52 & -5.69 & 4.27 & 0.95 & 0.59543\\
600 & c500 & -0.66 & -8.56 & 6.99 & 0.95 & 0.51694\\
240 & c1000 & 3.03 & 1.37 & 4.82 & 0.95 & 20.63183\\
280 & c1000 & 1.63 & 0.68 & 2.63 & 0.95 & 5.12611\\
320 & c1000 & 0.64 & -0.02 & 1.24 & 0.95 & 1.90342\\
360 & c1000 & -0.01 & -0.57 & 0.52 & 0.95 & 0.99277\\
400 & c1000 & -0.38 & -1.01 & 0.22 & 0.95 & 0.68359\\
440 & c1000 & -0.54 & -1.52 & 0.32 & 0.95 & 0.58403\\
480 & c1000 & -0.54 & -2.20 & 1.11 & 0.95 & 0.58190\\
520 & c1000 & -0.45 & -3.40 & 2.35 & 0.95 & 0.63546\\
560 & c1000 & -0.34 & -5.78 & 3.90 & 0.95 & 0.71487\\
600 & c1000 & -0.25 & -8.34 & 6.73 & 0.95 & 0.77863\\
240 & c1500 & 3.05 & 1.07 & 5.02 & 0.95 & 21.02227\\
280 & c1500 & 1.54 & 0.40 & 2.65 & 0.95 & 4.65584\\
320 & c1500 & 0.42 & -0.36 & 1.13 & 0.95 & 1.51502\\
360 & c1500 & -0.38 & -1.05 & 0.21 & 0.95 & 0.68077\\
400 & c1500 & -0.92 & -1.70 & -0.24 & 0.95 & 0.39703\\
440 & c1500 & -1.26 & -2.41 & -0.18 & 0.95 & 0.28245\\
480 & c1500 & -1.47 & -3.36 & 0.43 & 0.95 & 0.23037\\
520 & c1500 & -1.60 & -4.86 & 1.58 & 0.95 & 0.20247\\
560 & c1500 & -1.71 & -7.01 & 3.37 & 0.95 & 0.18021\\
600 & c1500 & -1.88 & -10.07 & 5.98 & 0.95 & 0.15267\\
240 & i500 & 2.31 & 0.79 & 3.93 & 0.95 & 10.10461\\
280 & i500 & 0.55 & -0.46 & 1.52 & 0.95 & 1.72468\\
320 & i500 & -0.54 & -1.13 & 0.08 & 0.95 & 0.58233\\
360 & i500 & -1.08 & -1.50 & -0.61 & 0.95 & 0.33902\\
400 & i500 & -1.22 & -1.78 & -0.65 & 0.95 & 0.29661\\
440 & i500 & -1.08 & -2.03 & -0.19 & 0.95 & 0.33991\\
480 & i500 & -0.81 & -2.16 & 0.59 & 0.95 & 0.44474\\
520 & i500 & -0.55 & -2.50 & 1.42 & 0.95 & 0.57904\\
560 & i500 & -0.42 & -3.16 & 2.28 & 0.95 & 0.65388\\
600 & i500 & -0.58 & -4.35 & 3.10 & 0.95 & 0.55820\\
240 & i1000 & 3.12 & 1.66 & 4.58 & 0.95 & 22.68463\\
280 & i1000 & 1.06 & 0.15 & 1.95 & 0.95 & 2.88377\\
320 & i1000 & -0.24 & -0.78 & 0.31 & 0.95 & 0.78490\\
360 & i1000 & -0.92 & -1.30 & -0.52 & 0.95 & 0.39866\\
400 & i1000 & -1.11 & -1.61 & -0.59 & 0.95 & 0.32935\\
440 & i1000 & -0.95 & -1.80 & -0.12 & 0.95 & 0.38574\\
480 & i1000 & -0.58 & -1.86 & 0.70 & 0.95 & 0.55825\\
520 & i1000 & -0.14 & -1.90 & 1.77 & 0.95 & 0.87013\\
560 & i1000 & 0.24 & -2.33 & 2.75 & 0.95 & 1.27313\\
600 & i1000 & 0.42 & -3.17 & 3.85 & 0.95 & 1.52411\\
240 & i1500 & 3.30 & 1.63 & 4.98 & 0.95 & 27.07329\\
280 & i1500 & 1.08 & 0.05 & 2.14 & 0.95 & 2.93821\\
320 & i1500 & -0.33 & -0.94 & 0.36 & 0.95 & 0.71847\\
360 & i1500 & -1.06 & -1.52 & -0.57 & 0.95 & 0.34502\\
400 & i1500 & -1.26 & -1.88 & -0.65 & 0.95 & 0.28360\\
440 & i1500 & -1.06 & -1.99 & -0.09 & 0.95 & 0.34778\\
480 & i1500 & -0.59 & -2.01 & 0.88 & 0.95 & 0.55459\\
520 & i1500 & 0.00 & -2.05 & 2.09 & 0.95 & 1.00234\\
560 & i1500 & 0.58 & -2.23 & 3.54 & 0.95 & 1.78962\\
600 & i1500 & 1.01 & -3.02 & 4.86 & 0.95 & 2.75123\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

@@

In the middle of the experiment (trial 1000), congruent and incongruent primes have positive effects in bins (200,240{]} and (240,280{]}, while incongruent primes again have negative effects in bins (320,360{]}, (360,400{]}, and (400,440{]}. Due to practicing stimulus-response associations, the primes generate a higher hazard of response occurrence for 80 ms (compared to 40 ms at the beginning of the experiment).

@@

240 & c500 & 17.02 & 6.26 & 29.22 & 0.95 & 24618458.6117376089096\\
280 & c500 & 8.49 & 3.71 & 13.54 & 0.95 & 4846.0911175084821\\
320 & c500 & 3.91 & 1.88 & 6.05 & 0.95 & 49.7913398591541\\
360 & c500 & 2.19 & 0.69 & 3.75 & 0.95 & 8.8918515266144\\
400 & c500 & 2.22 & -0.25 & 4.75 & 0.95 & 9.1907589464532\\
440 & c500 & 2.91 & -2.56 & 8.54 & 0.95 & 18.3094780953354\\
480 & c500 & 3.15 & -8.77 & 15.10 & 0.95 & 23.4104953301397\\
520 & c500 & 1.86 & -20.13 & 26.73 & 0.95 & 6.3973471906322\\
560 & c500 & -2.08 & -39.94 & 42.41 & 0.95 & 0.1244195286526\\
600 & c500 & -9.77 & -73.17 & 61.54 & 0.95 & 0.0000573483764\\
240 & c1000 & 14.46 & 4.94 & 26.35 & 0.95 & 1899836.0176862408407\\
280 & c1000 & 7.58 & 3.21 & 12.18 & 0.95 & 1961.8278183493471\\
320 & c1000 & 3.80 & 1.90 & 5.71 & 0.95 & 44.8743225515544\\
360 & c1000 & 2.02 & 0.72 & 3.35 & 0.95 & 7.5713234449885\\
400 & c1000 & 1.14 & -0.99 & 3.11 & 0.95 & 3.1378140505531\\
440 & c1000 & 0.06 & -4.41 & 4.87 & 0.95 & 1.0636745027736\\
480 & c1000 & -2.32 & -12.62 & 8.61 & 0.95 & 0.0982112115060\\
520 & c1000 & -7.10 & -27.24 & 15.97 & 0.95 & 0.0008224865296\\
560 & c1000 & -15.39 & -54.71 & 23.54 & 0.95 & 0.0000002080447\\
600 & c1000 & -28.27 & -92.96 & 35.54 & 0.95 & 0.0000000000005\\
240 & c1500 & 16.12 & 3.74 & 29.48 & 0.95 & 10001085.3946626689285\\
280 & c1500 & 8.54 & 2.78 & 14.43 & 0.95 & 5124.4351045810508\\
320 & c1500 & 4.22 & 1.75 & 6.70 & 0.95 & 68.1208056719069\\
360 & c1500 & 2.06 & 0.48 & 3.71 & 0.95 & 7.8233561607888\\
400 & c1500 & 0.95 & -1.75 & 3.26 & 0.95 & 2.5848218057173\\
440 & c1500 & -0.20 & -6.03 & 5.65 & 0.95 & 0.8181575226781\\
480 & c1500 & -2.49 & -15.23 & 10.07 & 0.95 & 0.0826145785621\\
520 & c1500 & -7.03 & -30.41 & 18.55 & 0.95 & 0.0008862056864\\
560 & c1500 & -14.91 & -58.81 & 27.21 & 0.95 & 0.0000003362909\\
600 & c1500 & -27.22 & -95.59 & 43.50 & 0.95 & 0.0000000000015\\
240 & i500 & -23.34 & -44.42 & -4.87 & 0.95 & 0.0000000000730\\
280 & i500 & -10.55 & -17.93 & -3.94 & 0.95 & 0.0000261643171\\
320 & i500 & -3.71 & -6.06 & -1.51 & 0.95 & 0.0246001196011\\
360 & i500 & -0.97 & -2.57 & 0.56 & 0.95 & 0.3780876098423\\
400 & i500 & -0.52 & -2.75 & 1.55 & 0.95 & 0.5916606671310\\
440 & i500 & -0.53 & -6.67 & 5.86 & 0.95 & 0.5871858652992\\
480 & i500 & 0.83 & -16.41 & 20.71 & 0.95 & 2.3019651321687\\
520 & i500 & 5.40 & -32.44 & 52.48 & 0.95 & 222.0443659277399\\
560 & i500 & 15.00 & -58.75 & 104.35 & 0.95 & 3282435.9279020344839\\
600 & i500 & 31.47 & -90.20 & 190.08 & 0.95 & 46319712352328.7578125000000\\
240 & i1000 & -21.85 & -43.05 & -4.10 & 0.95 & 0.0000000003243\\
280 & i1000 & -9.67 & -16.56 & -3.46 & 0.95 & 0.0000632158160\\
320 & i1000 & -3.17 & -5.23 & -0.99 & 0.95 & 0.0419655563481\\
360 & i1000 & -0.53 & -2.03 & 0.89 & 0.95 & 0.5909004105316\\
400 & i1000 & 0.09 & -1.88 & 2.11 & 0.95 & 1.0992267336787\\
440 & i1000 & 0.52 & -5.54 & 6.73 & 0.95 & 1.6827111411806\\
480 & i1000 & 2.58 & -14.16 & 22.53 & 0.95 & 13.2031868705690\\
520 & i1000 & 8.10 & -28.51 & 55.88 & 0.95 & 3307.4439707159477\\
560 & i1000 & 18.92 & -51.75 & 111.96 & 0.95 & 164758701.8390493392944\\
600 & i1000 & 36.86 & -89.39 & 191.12 & 0.95 & 10165856639901592.0000000000000\\
240 & i1500 & -20.51 & -42.95 & -2.49 & 0.95 & 0.0000000012362\\
280 & i1500 & -9.04 & -16.80 & -2.03 & 0.95 & 0.0001189822174\\
320 & i1500 & -2.86 & -5.47 & -0.25 & 0.95 & 0.0575421441866\\
360 & i1500 & -0.14 & -1.81 & 1.67 & 0.95 & 0.8709638702927\\
400 & i1500 & 0.94 & -1.61 & 3.40 & 0.95 & 2.5699339941536\\
440 & i1500 & 2.22 & -4.97 & 9.63 & 0.95 & 9.2076572160001\\
